Analysis Plan: Replicating Growth and Inequality in Public Good Provision

### 2022-10-08 analysis/reports/rmd/01_dataGMTV.Rmd
post data collection
added "quarto" and "paper" (already deprecated) + no clue what happened otherwise. Its been a long time

In an attempt to incorporate uncertainty to @GACHTER2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The first experiment will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty/tree/July21Replication) to visit the corresponding Github repository.

@@

# calculate gini based on stock (i.e. end of period earnings)
noPunish10[, stock0 := tokens - putin + sum/4]
noPunish10[, stock1 := other1 - pu1 + sum/4]
noPunish10[, stock2 := other2 - pu2 + sum/4]
noPunish10[, stock3 := other3 - pu3 + sum/4]

@@

>>>>

In an attempt to incorporate uncertainty to @GMTV2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The first experiment will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty/tree/July21Replication) to visit the corresponding Github repository.

@@

# calculate gini based on stock (i.e. end of period earnings)
noPunish10[, stock0 := (tokens - putin + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock1 := (other1 - pu1 + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock2 := (other2 - pu2 + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock3 := (other3 - pu3 + 1.5*sum/4) %>% ceiling()]

@@

Unfortunately, I made a misstake calculating the ```share```. For this reason, it will be recalculated here:

```{r recalcShare}
GMTV[, share := contribution/endowment]
```

### 2022-10-08 analysis/reports/rmd/02_dataProcessing.Rmd
added "quarto" and "paper" (already deprecated) + no clue what happened otherwise. Its been a long time

In an attempt to incorporate uncertainty to @GACHTER2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The first experiment will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

>>>>

In an attempt to incorporate uncertainty to @GMTV2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The first experiment will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

### 2022-10-08 analysis/reports/rmd/04_prettyAnalyses.Rmd
added "quarto" and "paper" (already deprecated) + no clue what happened otherwise. Its been a long time

---
title: "3. Analysis Plan"
description: |
  Analyze and compare data
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: ../biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
---

```{r setup, results = "hide", echo = FALSE}
if(knitr::is_html_output()){
  knitr::opts_chunk$set(echo = FALSE)
}else{
  knitr::opts_chunk$set(echo = TRUE)
}

@@

```{r packages}
library(data.table)
library(magrittr)
library(kableExtra)
library(downloadthis)
library(ggplot2)
library(patchwork)
library(stargazer)
```

@@

# Key Takeaways

- We replicate a public goods experiment with dynamic interdependency and find similar results as @GACHTER2017.
- Absolute contributions increase over time.
- Just as in _static_ public goods experiments, the share of endowments contributed decreases over time.
- The richest groups earn fifteen times more than the poorest groups.
- While there clearly is growth, groups do not realize the maximal potential efficiency and earn just over `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`% of what is possible.
- Even though we conducted the experiment online and remotely, dropouts (or "attrition") are no concern.
- Relying on an inexperienced non-convenience sample, `r trunc((R1[comprehension > 1] %>% NROW() / R1[!(is.na(comprehension))] %>% NROW())*100)`% of all participants stated that they understood the game.
- Participants make relatively fast decisions which makes longer games feasible in the future.^[Taken together with inefficient growth and the lack of dropouts, additional rounds are fairly cheap.]

@@

In an attempt to incorporate uncertainty to @GACHTER2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. This experiment  replicated Gächter et al.'s NOPUNISH 10-round treatment arm as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

@@

I'll refer to the data @GACHTER2017 provided as _GMTV_ or _noPunish10_ in what follows. GMTV conducted most of their sessions in late 2012. All of these sessions ran in Nottingham using a student sample. collected They `r GMTV[, groupID %>% unique() %>% length()]` observations in their NOPUNISH 10 treatment arm. Our data, refered to as _replication_ is more recent, gathered remotely in Hamburg using a different tech stack as well as a different sample.

@@

```{r meta}
meta %>% kable()
```

@@

### Hm..?

The plot above puzzles me a little as the right panel visualizes the mean `share==contribution/endowment` by treatment and round. This yields a different result than the mean of the sum of contributions divided by the sum of endowments per treatment and round as the following table illustrates.

```{r}
main[round == 10, 
     .(contribution = mean(contribution), 
       endowment = mean(endowment), 
       # share = mean(share), # same thing as the following line
       share = mean(contribution/endowment),
       sumContBySumEndo = mean(sum(contribution)/sum(endowment))), 
     by = c("treatment", "round")] %>% 
  kable()
```

```{r plotContributionsStudents, eval = FALSE}
SUM <- main[treatment == "replication",
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "student"),
            .SDcols = "contribution"]

p1 <- ggplot(data = SUM, 
             aes(x = round, y = contribution, fill = student, color = student, lty = student)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Average Amount of Tokens contributed") +
  scale_color_manual(values = colors) + 
  theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

@@

Possibly of more interest are the implications contributions have for wealth generation and growth. To measure growth, we define a variable _stock_ which sums the endowments of all participants in a given group at the end of the round (that is, after the contributions have been made, multiplied and redistributed). @GACHTER2017 refer to that variable as "wealth" so we will do the same in what follows. Before the start of round 1, wealth will be 80 in all groups by construction. The maximal wealth that can be reached in round 10 (if everyone contributes their entire endowment in each round) is approximately `r (80*1.5^10) %>% trunc()` tokens or `r (80*1.5^10 /20) %>% trunc()` Euro per group. Table 3 shows some summary statistics regarding wealth. Groups do achieve growth on average. While there is clearly growth, groups do not realize the maximal potential efficiency as the replication groups reach on average a level of `r main[treatment == "replication" & round == 10, mean(stock)] %>% trunc()` tokens out of `r (80*1.5^10) %>% trunc()` maximally possible or `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`%. As in the original data, there is large heterogeneity with the richest group reaching `r main[treatment == "replication" & round == 10, max(stock)]` tokens whereas the poorest group ends up with `r main[treatment == "replication" & round == 10, min(stock)]` tokens.

@@

We next consider whether we were able to replicate GMTV's results with our data. Absolute contributions tended to be higher in @GACHTER2017 but end up at around the same level as in the replication due to a stark decline in contributions on the last round. In terms of shares contributed both data sources exhibit a similar pattern: they decline and do not stabilize. Even though the share of current endowments contributed in the last round is quite similar, the share declined a little faster in our data.

@@

```{r summary3, results = "asis"}

@@

```{r}
main[treatment == "replication" & round == 10, 
     lapply(.SD, mean, na.rm = TRUE), 
     by = c("round", "student"), 
     .SDcols = c("contribution", "share", "stock")] %>%
  kable()
```

>>>>

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ../prettyReports/svm.tex
bibliography: ../biblio.bib
header-includes:
  -  \usepackage{hyperref}
  - \usepackage{array}   
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[table]{xcolor}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{fontawesome}
  - \usepackage[para,online,flushleft]{threeparttable}
biblio-style: apsr
title: "Analysis Plan: Replicating Growth and Inequality in Public Good Provision"
thanks: "Further replication files are available on the author's Github account (http://github.com/howquez). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: roggenkh@hsu-hh.de."
author:
- name: Hauke Roggenkamp
  affiliation: Helmut-Schmidt-University
- name: Stefan Traub
  affiliation: Helmut-Schmidt-University
- name: Michael Berlemann
  affiliation: Helmut-Schmidt-University
abstract: "We replicate a public goods experiment with dynamic interdependency and find similar results as GMTV. Absolute contributions increase over time. Just as in static public goods experiments, the share of endowments contributed decreases over time. The richest groups earn fifteen times more than the poorest groups. While there clearly is growth, groups do not realize the maximal potential efficiency and earn just a fraction of what is possible. Even though we conducted the experiment online and remotely, dropouts are no concern. Participants make relatively fast decisions which makes longer games feasible in the future. Taken together with inefficient growth and the lack of dropouts, additional rounds are fairly cheap."
keywords: "dynamic public goods game, inequality, cooperation, growth"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: libertine
fontsize: 11pt
# spacing: double
endnote: no
---

```{r setup, results = "hide", echo = FALSE}
if(knitr::is_html_output()){
  knitr::opts_chunk$set(echo = FALSE)
  evalKable <- TRUE
}else{
  knitr::opts_chunk$set(echo = FALSE)
  evalKable <- FALSE
}

@@

```{r packages}
library(data.table)
library(magrittr)
library(kableExtra)
library(downloadthis)
library(ggplot2)
library(patchwork)
library(stargazer)
library(Rmisc)
```

@@

@@

In an attempt to incorporate uncertainty to @GMTV2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. This experiment  replicated Gächter et al.'s NOPUNISH 10-round treatment arm as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

@@

I'll refer to the data @GMTV2017 provided as _GMTV_ or _noPunish10_ in what follows. GMTV conducted most of their sessions in late 2012. All of these sessions ran in Nottingham using a student sample. collected They `r GMTV[, groupID %>% unique() %>% length()]` observations in their NOPUNISH 10 treatment arm. Our data, refered to as _replication_ is more recent, gathered remotely in Hamburg using a different tech stack as well as a different sample.

@@

```{r meta, eval = evalKable}
meta %>% kable()
```

@@

@@

Possibly of more interest are the implications contributions have for wealth generation and growth. To measure growth, we define a variable _stock_ which sums the endowments of all participants in a given group at the end of the round (that is, after the contributions have been made, multiplied and redistributed). @GMTV2017 refer to that variable as "wealth" so we will do the same in what follows. Before the start of round 1, wealth will be 80 in all groups by construction. The maximal wealth that can be reached in round 10 (if everyone contributes their entire endowment in each round) is approximately `r (80*1.5^10) %>% trunc()` tokens or `r (80*1.5^10 /20) %>% trunc()` Euro per group. Table 3 shows some summary statistics regarding wealth. Groups do achieve growth on average. While there is clearly growth, groups do not realize the maximal potential efficiency as the replication groups reach on average a level of `r main[treatment == "replication" & round == 10, mean(stock)] %>% trunc()` tokens out of `r (80*1.5^10) %>% trunc()` maximally possible or `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`%. As in the original data, there is large heterogeneity with the richest group reaching `r main[treatment == "replication" & round == 10, max(stock)]` tokens whereas the poorest group ends up with `r main[treatment == "replication" & round == 10, min(stock)]` tokens.

@@

We next consider whether we were able to replicate GMTV's results with our data. Absolute contributions tended to be higher in @GMTV2017 but end up at around the same level as in the replication due to a stark decline in contributions on the last round. In terms of shares contributed both data sources exhibit a similar pattern: they decline and do not stabilize. Even though the share of current endowments contributed in the last round is quite similar, the share declined a little faster in our data.

@@

```{r summary3, results = "asis", warning = FALSE}

@@

```{r eval = evalKable}
main[treatment == "replication" & round == 10, 
     lapply(.SD, mean, na.rm = TRUE), 
     by = c("round", "student"), 
     .SDcols = c("contribution", "share", "stock")] %>%
  kable()
```
